Page not found

### 2023-06-21 manuscript/chapters/2_neural_fields.qmd
plop

### Gaussian $\mathcal{N}(\vec 0, \mathbb{1}_N)$ low-rank network of neurons

@@

::: {.content-visible when-format="pdf"}
![(TODO) Simulation of the Gaussian network of neurons with $p=1$. Every neuron is represented by a dot, of which the height represents its activity level, and the colorbar doubles as a y-axis.](figures/lol7.jpg){#fig-p1}
:::

@@

::: {.content-visible when-format="pdf"}
![(TODO PDF) Simulation of the Gaussian network of neurons with $p=2$. Every neuron is represented by a dot, of which the height represents its activity level. Projections are shown by gray points.](figures/lol7.jpg){width=500px #fig-p2}
:::

@@

::: {.content-visible when-format="pdf"}
![(TODO PDF) A cycling neural field with $\delta=6$ and initial condition $h(z_1, z_2, t < 0) = z_1$. The latent trajectory is also computed, and demonstrates the oscillatory behaviour.](figures/lol7.jpg){#fig-cycling-anim}
:::

>>>>

### Gaussian low-rank network of neurons

@@

<!-- :::: {.content-visible when-format="pdf" layout-nrow=2}
:::{#fig-p1 layout-nrow=2}
![First frame](figures/embedding_p=1.mp4-1.jpg)

![Last frame](figures/embedding_p=1.mp4-2.jpg)

Simulation of the Gaussian network of neurons with $p=1$. Every neuron is represented by a dot, of which the height represents its activity level, and the colorbar doubles as a y-axis. The first and the last frame of the animation are shown, [available here](https://ninivert.github.io/lcnthesis/chapters/figures/embedding_p=1.mp4).
:::
:::: -->

::: {.content-visible when-format="pdf"}
```{python}
#| label: fig-p1
#| layout-nrow: 2
#| fig-cap: Simulation of the Gaussian network of neurons with $p=1$. Every neuron is represented by a dot, of which the height represents its activity level, and the colorbar doubles as a y-axis. The first and the last frame of the animation are shown, [available here](https://ninivert.github.io/lcnthesis/chapters/figures/embedding_p=1.mp4).
#| fig-subcap:
#|   - First frame
#|   - Last frame
#| echo: false
# Note: I have to do this wizardry because Quarto *really* doesn't seem to like nested divs with conditional content...
from IPython.display import Image
display(Image('figures/embedding_p=1.mp4-1.jpg'))
display(Image('figures/embedding_p=1.mp4-2.jpg'))
```
:::

@@

::: {.content-visible when-format="pdf"}
```{python}
#| label: fig-p2
#| layout-ncol: 2
#| fig-cap: Simulation of the Gaussian network of neurons with $p=2$. Every neuron is represented by a dot, of which the height represents its activity level. Projections are shown by gray points. The first and the last frame of the animation are shown, [available here](https://ninivert.github.io/lcnthesis/chapters/figures/embedding_p=2.mp4).
#| fig-subcap:
#|   - First frame
#|   - Last frame
#| echo: false
# Note: I have to do this wizardry because Quarto *really* doesn't seem to like nested divs with conditional content...
from IPython.display import Image
display(Image('figures/embedding_p=2.mp4-1.jpg'))
display(Image('figures/embedding_p=2.mp4-2.jpg'))
```
:::

@@

::: {.content-visible when-format="pdf"}
```{python}
#| label: fig-cycling-anim
#| layout-nrow: 2
#| fig-cap: A cycling neural field with $\delta=6$ and initial condition $h(z_1, z_2, t < 0) = z_1$. The latent trajectory is also computed, and demonstrates the oscillatory behaviour. The first and the last frame of the animation are shown, [available here](https://ninivert.github.io/lcnthesis/chapters/figures/embedding_p=2_cycling.mp4).
#| fig-subcap:
#|   - First frame
#|   - Last frame
#| echo: false
# Note: I have to do this wizardry because Quarto *really* doesn't seem to like nested divs with conditional content...
from IPython.display import Image
display(Image('figures/embedding_p=2_cycling.mp4-1.jpg'))
display(Image('figures/embedding_p=2_cycling.mp4-2.jpg'))
```
:::

### 2023-06-21 manuscript/chapters/3_mappings.qmd
plop

[^indices]: In the source code, only the method `indices(v12: array[float, 2]) -> array[int, 1]` is defined. This is because some mappings are implemented (for instance, the [`RecursiveLocalMapping`](https://github.com/ninivert/lcnthesis/blob/master/notebooks/neurodyn/_mapping.py#L244), that we don't discuss here), which don't lend themselves to the simple permutation of bits formulation.

@@

* @fig-cg-2d-1d-random: The random mapping can be numerically evaluated as a trivial case of a mapping that does not have locality. The coarse-graining averages out all the structure of the activity, and the recurrent currents are zero, because the contributions from excitatory and inhibitory populations cancel out. This results in the erasure of the cycling behavior in the latent trajectory.
* @fig-cg-2d-1d-column: The Column mapping shows the projection of the initial condition onto the $z_1$ axis, yielding the sigmoid in the $[0,1]$ embedding. Again, the current currents cancel out and yield zero, and the latent trajectory quickly decays towards $(\kappa_1, \kappa_2) = (0,0)$.
* @fig-cg-2d-1d-z: The Z-mapping shows that despite the fractal structure of the activity, there is still regularity in the $[0,1]$ embedding, resulting in non-zero recurrent currents. In consequence, the latent trajectory shows a clear cycling behavior, that closely matches that of the 2D neural field in @fig-cg-2d-1d-original.

@@

::: {#fig-cg-2d-1d}

@@

The cycling 2D neural field ($4^8$ square populations) is mapped to $[0,1]$ using different mappings, then coarse-grained (resulting in $2^8$ segment populations).
:::

@@

::: {.content-visible when-format="pdf"}
![(TODO PDF. initial state and final state) Equivalence of dynamics, from 3D to 2D to 1D.](figures/lol7.jpg){#fig-cg-3d-2d-1d}
:::

>>>>

[^indices]: In the source code, only the method `indices(v12: array[float, 2]) -> array[int, 1]` is defined. This is because some mappings are implemented (for instance, the [`RecursiveLocalMapping`](https://github.com/ninivert/lcnthesis/blob/master/notebooks/neurodyn/_mapping.py#L244), that we don't discuss here), that don't lend themselves to the simple permutation of bits formulation.

@@

* The random mapping can be numerically evaluated as a trivial case of a mapping that does not have locality. The coarse-graining averages out all the structure of the activity, and the recurrent currents are zero, because the contributions from excitatory and inhibitory populations cancel out. This results in the erasure of the cycling behavior in the latent trajectory.
* The Column mapping shows the projection of the initial condition onto the $z_1$ axis, yielding the sigmoid in the $[0,1]$ embedding. Again, the current currents cancel out and yield zero, and the latent trajectory quickly decays towards $(\kappa_1, \kappa_2) = (0,0)$.
* The Z-mapping shows that despite the fractal structure of the activity, there is still regularity in the $[0,1]$ embedding, resulting in non-zero recurrent currents. In consequence, the latent trajectory shows a clear cycling behavior, that closely matches that of the 2D neural field.

@@

::: {.content-visible when-format="html"}

:::::: {#fig-cg-2d-1d}

@@

The cycling 2D neural field ($4^8$ square populations) is mapped to $[0,1]$ using different mappings, then coarse-grained (resulting in $2^8$ segment populations).
::::::

:::

::: {.content-visible when-format="pdf"}
```{python}
#| label: fig-cg-2d-1d
#| layout-ncol: 2
#| layout-nrow: 4
#| fig-cap: The cycling 2D neural field ($4^8$ square populations) is mapped to $[0,1]$ using different mappings, then coarse-grained (resulting in $2^8$ segment populations).
#| fig-subcap:
#|   - Neural field in the 2D embedding. First frame, animation [available here](https://ninivert.github.io/lcnthesis/chapters/figures/coarse_graining_p=2_original.mp4).
#|   - Neural field in the 2D embedding. Last frame, animation [available here](https://ninivert.github.io/lcnthesis/chapters/figures/coarse_graining_p=2_original.mp4).
#|   - Random mapping followed by coarse-graining. First frame, animation [available here](https://ninivert.github.io/lcnthesis/chapters/figures/coarse_graining_p=2_random.mp4).
#|   - Random mapping followed by coarse-graining. Last frame, animation [available here](https://ninivert.github.io/lcnthesis/chapters/figures/coarse_graining_p=2_random.mp4).
#|   - Column mapping followed by coarse-graining. First frame, animation [available here](https://ninivert.github.io/lcnthesis/chapters/figures/coarse_graining_p=2_column.mp4).
#|   - Column mapping followed by coarse-graining. Last frame, animation [available here](https://ninivert.github.io/lcnthesis/chapters/figures/coarse_graining_p=2_column.mp4).
#|   - Z-mapping followed by coarse-graining. First frame, animation [available here](https://ninivert.github.io/lcnthesis/chapters/figures/coarse_graining_p=2_z.mp4).
#|   - Z-mapping followed by coarse-graining. Last frame, animation [available here](https://ninivert.github.io/lcnthesis/chapters/figures/coarse_graining_p=2_z.mp4).
#| echo: false
# Note: I have to do this wizardry because Quarto *really* doesn't seem to like nested divs with conditional content...
from IPython.display import Image
display(Image('figures/coarse_graining_p=2_original.mp4-1.jpg'))
display(Image('figures/coarse_graining_p=2_original.mp4-2.jpg'))
display(Image('figures/coarse_graining_p=2_random.mp4-1.jpg'))
display(Image('figures/coarse_graining_p=2_random.mp4-2.jpg'))
display(Image('figures/coarse_graining_p=2_column.mp4-1.jpg'))
display(Image('figures/coarse_graining_p=2_column.mp4-2.jpg'))
display(Image('figures/coarse_graining_p=2_z.mp4-1.jpg'))
display(Image('figures/coarse_graining_p=2_z.mp4-2.jpg'))
```
:::

@@

::: {.content-visible when-format="pdf"}
```{python}
#| label: fig-cg-3d-2d-1d
#| layout-nrow: 2
#| fig-cap: Equivalence of dynamics, from 3D to 2D to 1D. The first and the last frame of the animation are shown, [available here](https://ninivert.github.io/lcnthesis/chapters/figures/embedding_3d_2d_1d.mp4).
#| fig-subcap:
#|   - First frame
#|   - Last frame
#| echo: false
# Note: I have to do this wizardry because Quarto *really* doesn't seem to like nested divs with conditional content...
from IPython.display import Image
display(Image('figures/embedding_3d_2d_1d.mp4-1.jpg'))
display(Image('figures/embedding_3d_2d_1d.mp4-2.jpg'))
```
:::

### 2023-06-21 manuscript/chapters/5_conclusion.qmd
plop

In this thesis, we have constructed an example of neural field dynamics in $\R^2$, that can be mapped to a neural field equation in $[0,1]$, via well-known measurable bijections (such as the Z-mapping) between $[0,1]^2$ and $[0,1]$. To illustrate this, we have introduced a *coarse-graining* procedure in which numerical simulations of $2^n$ segment populations in $[0,1]$ converge to the same dynamics obtained from simulations of $4^n$ square populations in $[0,1]^2$. Additionally, we showed that the reduction of dimensionality using the mappings can be repeated, even if the starting neural field has a fractal structure. To help quantify how the mapping conserves the regularity between neural fields, we introduced a *measure of locality*, based on "average variation". Locality guarantees that the neural field equation on $[0,1]$ can be numerically approximated using classical grid-based methods.

This thesis has been illustrated with simulations of low-rank networks, however, to demonstrate the generality of our result, future work would try to numerically study an example of neural field dynamics that does not have a low-rank structure. This would also avoid confusion between the linear dimensionality of the dynamics and the dimensionality of the embedding, which are the same in the context of low-rank networks.
While we have presented some mathematical arguments for (1) the equivalence of the neural field dynamics in 2D and 1D, and (2) the convergence of the numerical approximation of the 1D neural field dynamics using locality, we could use these arguments to make rigorous proofs of both statements.

The use of the notion of regularity for the analysis of large-scale neuronal recordings is in its infancy @Stringer2019. This purely theoretical thesis gives a concrete example of an important conceptual fact: if we want to analyse large-scale neuronal recordings via nonlinear dimensionality reduction techniques (e.g. if we want to model neuronal activity as neural fields evolving on some abstract latent embedding space) the notion of dimensionality is meaningless without the notion of regularity. Indeed, our example shows that some neural field dynamics (neuronal population dynamics) can be equivalently expressed as dynamics on the embedding space $\R^2$ and the embedding space $[0,1]$ (but the regularity of the corresponding neural fields are different).

>>>>

In this thesis, we have constructed an example of neural field dynamics in $\R^2$, that can be mapped to a neural field equation in $[0,1]$, via well-known measurable bijections (such as the Z-mapping) between $[0,1]^2$ and $[0,1]$. To illustrate this, we have introduced a *coarse-graining* procedure in which numerical simulations of $2^n$ segment populations in $[0,1]$ converge to the same dynamics obtained from simulations of $4^n$ square populations in $[0,1]^2$. Additionally, we showed that the reduction of dimensionality using the mappings can be repeated, even if the starting neural field has a fractal structure. To help quantify how the mapping conserves the regularity between neural fields, we introduced a measure of locality, based on "average variation". Locality guarantees that the neural field equation on $[0,1]$ can be numerically approximated using classical grid-based methods.

Although, in this thesis, we have used simulations of simple low-rank networks for numerical illustrations, to demonstrate the generality of our result, future work would try to numerically study an example of neural field dynamics that does not have a low-rank structure. This would also avoid confusion between the linear dimensionality of the dynamics and the dimensionality of the embedding, which are the same in the context of low-rank networks.
While we have presented some mathematical arguments for (1) the equivalence of the neural field dynamics in 2D and 1D, and (2) the convergence of the numerical approximation of the 1D neural field dynamics using locality, we could use these arguments to make rigorous proofs of both statements.

The use of the notion of regularity for the analysis of large-scale neuronal recordings is in its infancy @Stringer2019. This purely theoretical thesis gives a concrete example of an important conceptual fact: if we want to analyse large-scale neuronal recordings via nonlinear dimensionality reduction techniques (e.g. if we want to model neuronal activity as neural fields evolving on some abstract latent embedding space) the notion of dimensionality is meaningless without an associated notion of regularity. Indeed, our example shows that some neural field dynamics (neuronal population dynamics) can be equivalently expressed as dynamics on the embedding space $\R^2$ and the embedding space $[0,1]$ (but the regularity of the corresponding neural fields are different).

### 2023-06-21 manuscript/index.qmd
plop

# Index {.unnumbered .hidden}

\[TODO : a nice illustration for this, maybe some cool animation ?\]

>>>>

::: {.content-visible when-format="html"}

::: {.content-hidden}
# Abstract {.unnumbered}
:::

:::

::: {.content-visible when-format="html"}
**Abstract**
:::

Early recordings of cat's somatosensory and visual cortices suggested that the cortical sheet is organized in vertical columns of functionally similar neurons.
The columnar organization of the cortex motivated the design of spatially structured models of neural population dynamics, namely neural field models, where the spatial dimensions corresponded to the two dimensions of the cortical sheet.
Later it has been shown that not only the physical position of the neurons can be used to parametrize their function, but also their location in more abstract "embedding spaces", for example a space of "neuron celltypes".

Decoupling the embedding space from the physical position of neurons on the cortical sheet begs the question of uniqueness of the embedding space. Given an embedding space and neural dynamics thereon, are there other embedding spaces which can express the same behavior? Is there a "natural" choice for the embedding space?

In this work, we show that the answer to the second question is highly nontrivial because given some neural dynamics, we can find multiple equivalent embedding spaces in different dimensions. We illustrate this using a toy example of a 2 (or 3)-dimensional embedding that is mapped to an equivalent 1-dimensional embedding. This is done by applying measurable bijective mappings $S : [0,1]^2 \mapsto [0,1]$, which lead, as we show, to well-behaved numerical simulation schemes. In this, we exploit the key concept of "locality", which expresses the property of (possibly fractal) neural fields, in which neighbouring neurons in the embedding space *tend* to have similar functions.

![Smooth neural fields in $p$ dimensions can be mapped to equivalent neural fields in one dimension, and the dynamics are equivalent.](chapters/figures/fig-cover.svg)
